---
title: "Reproducible Research Assignment 1"
author: "Scott Barker"
date: "March 5, 2016"
output: html_document
---
## Loading and preprocessing the data

##<<my-label, eval=TRUE, dev='png'>>=
  
1. Load the data(ie.. read.csv)  

```{r}
	if(file.exists("activity.csv")) {
  activity <- read.csv("activity.csv", colClass =c ('integer', 'Date', 'integer'))
}
```

2. Process/transform the data (if necessary) into a format suitable for your analysis
	
```{r}
  stepsPerday <- aggregate(steps ~ date, activity, sum)

    head(stepsPerday)
    stepsPerday
    summary(stepsPerday)
```
    
## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r}
mean(stepsPerday$steps)
```

2. Make a histogram of the total number of steps taken each day
```{r}
barplot(stepsPerday$steps, names.arg=stepsPerday$date, ylim=c(0, 30000), xlab = "Date", ylab="sum(steps)", )
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(stepsPerday$steps)

median(stepsPerday$steps)
```
## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps.interval <- aggregate(steps ~ interval, activity, mean)
plot(steps.interval, type='l')
```

2.	Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
steps.interval$interva[which.max(steps.interval$steps)]
```

## Imputing Data 

1.Calculate and report the total number of missing values in the dataset (i.e. the
```{r}
missingval <- sum(is.na(activity$steps))
missingval 
```
2. Devise a strategy for filling in all of the missing values in the dataset. 
Create a new dataset that is equal to the original dataset but with the missing data filled in.
	
```{r}
activity.clean <- merge(activity, stepsPerday, by="date", suffixes=c("", ".mean"))
  nas <- is.na(activity.clean$steps)
  activity.clean$steps[nas] <- activity.clean$steps.mean[nas]
  activity.clean <- activity.clean[, c(1:3)]
  head(activity.clean)
  summary(activity.clean)
  stepsclean <- aggregate(steps ~ date, activity.clean, sum)
  stepsclean
  
```

3.	Make a histogram of the total number of steps taken each day.

```{r}
stepsPerday <- aggregate(steps ~ date, activity.clean, sum)
barplot(stepsPerday$steps, names.arg=stepsPerday$date, ylim=c(0, 30000), 
        xlab="date", ylab="sum(steps)",)
```

4. Calculate and report the mean and median total number of steps taken per day.

```{r}
mean(stepsPerday$steps)

median(stepsPerday$steps)
```
No changes in the mean and median values of the data after the addition of the missing data because the steps were NA for the day.

## Are there differences in activity patterns between weekdays and weekends?

	1	Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
	
```{r}
dayType <- function(dates) {
  f <- function(date) {
    if (weekdays(date) %in% c("Saturday", "Sunday")) {
      "weekend"
    }
    else {
      "weekday"
    }
  }
  sapply(dates, f)
}
activity$dayType <- as.factor(dayType(activity$date))
str(activity)
```

2. Make a panel plot containing a time series plot (i.e. type ="l") of the 5-minute interval (x-axis) and the average numbrer of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
library(lattice)

steps.interval <- aggregate(steps ~ interval + dayType, activity, mean)
xyplot(steps ~ interval | dayType, data=steps.interval, layout=c(2,1), type='l')
```

